The \emph{FThresholds} package for \emph{Macaulay2}

### 2019-06-05 Fsing/FThresholdsPaper/FThresholds.tex
updates to section 4 of paper FThresholds

There are three main sections in this paper, which describe and illustrate the main functionality implemented in the package \emph{FThresholds}.
\begin{itemize}
\item \autoref{sec.Nu} describes the methods implemented for approximating the $F$-pure threshold and other $F$-thresholds.  
\item \autoref{sec.IsFPT} explains the package's functionality for determining whether a given number is the $F$-pure threshold, or more generally, an \emph{$F$-jumping exponent}.
\item \autoref{sec.FPT} summarizes the features of the function {\tt fpt} and how it works. 
\end{itemize}

@@

In more detail, let $e$ denote the value of the option {\tt DepthofSearch}, which conservatively defaults to $1$.  The {\tt fpt} function first computes $\nu=\nu_f(p^e)$, which agrees with the output of {\tt nu(e,f)}.  It is well-known that the $F$-pure threshold of $f$ is greater than $\nu/p^e$ and at most $(\nu+1)/p^e$, and applying  \cite[Proposition 4.2]{HernandezFPurityOfHypersurfaces} to this lower bound tells us that the $F$-pure threshold of $f$ must be at least $\nu/(p^e-1)$.  In summary, we know that the $F$-pure threshold of $f$ must lie in the closed interval 
%
\begin{equation}
\label{estimating-interval: e}
\tag{$\dagger$}
\left[ \frac{\nu}{p^e-1}, \frac{\nu+1}{p^e} \right].
\end{equation}

With these estimates in hand, the subroutine {\tt guessFPT} is called to make some ``educated guesses" in an attempt to identify the $F$-pure threshold within this interval, or at least narrow down this interval to produce improved estimates.  The number of ``guesses" is controlled by the option {\tt Attempts}, which conservatively defaults to 3.  If {\tt Attempts} is set to 0, then {\tt guessFPT} is bypassed. If  {\tt Attempts} is set to at least 1, then a first check is run to verify whether the right-hand endpoint $(\nu+1)/p^e$ of the above interval \eqref{estimating-interval: e} the $F$-pure threshold.  We illustrate this below.

@@

If  {\tt Attempts} is set to at least 2, and the right-hand endpoint $(\nu+1)/p^e$ of the interval \eqref{estimating-interval: e} is not the $F$-pure threshold, then a second check is run to verify whether the left-hand endpoint $\nu/(p^e-1)$ of this interval is the $F$-pure threshold.

@@

            If neither endpoint is the $F$-pure threshold, and {\tt Attempts} is set to more than 2, additional checks are performed at certain numbers within the interval .  A number in the interval \eqref{estimating-interval: e} with minimal denominator is selected, and {\tt compareFPT} is used to test that number. If that ``guess" is correct, its value is returned; otherwise, the information returned by {\tt compareFPT} is used to narrow down the interval, and this process is repeated as many times as specified by {\tt Attempts}.

@@

If {\tt guessFPT} is unsuccessful and {\tt UseFSignature} is set to {\tt true}, the fpt function proceeds to use the convexity of the $F$-signature function and a secant line argument to attempt to narrow down the interval bounding the $F$-pure threshold.

@@

When {\tt FRegularityCheck} is set to {\tt true} and no exact answer has been found, a final check is run (if necessary) to verify whether the final lower bound for the $F$-pure threshold is the exact answer.

@@

The computations performed when {\tt UseFSignature} and {\tt FRegularityCheck} are set to {\tt true} are often slow, and often fail to improve the estimate, and for this reason, these options should be used sparingly.
            It is often more effective to increase the values of {\tt Attempts} or {\tt DepthOfSearch}, instead.

@@

i31: f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

@@

As seen in several examples above, when the exact answer is not found, a list containing the endpoints of an interval containing the $F$-pure threshold of $f$ is returned.  Whether that interval is open, closed, or a mixed interval depends on the options passed; if the option {\tt Verbose} is set to {\tt true}, the precise interval will be printed.

>>>>

There are three main sections in this paper, which describe and illustrate the main functionality implemented in the package \emph{FThresholds}.
\begin{itemize}
\item \autoref{sec.Nu} describes the methods implemented for approximating the $F$-pure threshold and other $F$-thresholds.  
\item \autoref{sec.IsFPT} explains the package's functionality for determining whether a given number is the $F$-pure threshold, or more generally, an \emph{$F$-jumping exponent}.
\item \autoref{sec.FPT} describes the features of the central function of the package, {\tt fpt}. 
\end{itemize}

@@

In more detail, let $e$ denote the value of the option {\tt DepthOfSearch}, which conservatively defaults to {\tt 1}.  The {\tt fpt} function first computes $\nu=\nu_f(p^e)$, which agrees with the output of {\tt nu(e,f)}.  It is well known that the $F$-pure threshold of $f$ is greater than $\nu/p^e$ and at most $(\nu+1)/p^e$, and applying  \cite[Proposition 4.2]{HernandezFPurityOfHypersurfaces} to this lower bound tells us that the $F$-pure threshold of $f$ must be at least $\nu/(p^e-1)$.  In summary, we know that the $F$-pure threshold of $f$ must lie in the closed interval 
%
\begin{equation}
\label{estimating-interval: e}
\tag{$\dagger$}
\left[ \frac{\nu}{p^e-1}, \frac{\nu+1}{p^e} \right].
\end{equation}

With these estimates in hand, the subroutine {\tt guessFPT} is called to make some ``educated guesses" in an attempt to identify the $F$-pure threshold within this interval, or at least narrow down this interval to produce improved estimates.  The number of ``guesses" is controlled by the option {\tt Attempts}, which conservatively defaults to {\tt 3}.  If {\tt Attempts} is set to {\tt 0}, then {\tt guessFPT} is bypassed. If  {\tt Attempts} is set to at least {\tt 1}, then a first check is run to verify whether the right-hand endpoint $(\nu+1)/p^e$ of the above interval \eqref{estimating-interval: e} is the $F$-pure threshold.  We illustrate this below.

@@

If  {\tt Attempts} is set to at least {\tt 2} and the right-hand endpoint $(\nu+1)/p^e$ of the interval \eqref{estimating-interval: e} is not the $F$-pure threshold, then a second check is run to verify whether the left-hand endpoint $\nu/(p^e-1)$ of this interval is the $F$-pure threshold.

@@

            If neither endpoint is the $F$-pure threshold and {\tt Attempts} is set to more than {\tt 2}, then  additional checks are performed at certain numbers within the interval .  A number in the interval \eqref{estimating-interval: e} with minimal denominator is selected, and {\tt compareFPT} is used to test that number. If that ``guess" is correct, its value is returned; otherwise, the information returned by {\tt compareFPT} is used to narrow down the interval, and this process is repeated as many times as specified by {\tt Attempts}.

@@

If {\tt guessFPT} is unsuccessful and {\tt UseFSignature} is set to {\tt true}, then the {\tt fpt} function proceeds to use the convexity of the $F$-signature function and a secant line argument to attempt to narrow down the interval bounding the $F$-pure threshold.

@@

When {\tt FRegularityCheck} is set to {\tt true} and no exact answer has been found, a final check is run to verify whether the final lower bound for the $F$-pure threshold is the exact answer (if it has not already been eliminated as a possibility).

@@

The computations performed when {\tt UseFSignature} and {\tt FRegularityCheck} are set to {\tt true} are often slow, and often fail to improve the estimate, and for this reason, these options should be used sparingly.
            It is often more effective to increase the values of {\tt Attempts} or {\tt DepthOfSearch} instead.

@@

i31 : f = x^7*y^5*(x + y)^5*(x^2 + y^3)^4;

@@

As seen in several examples above, when the exact  $F$-pure threshold of $f$ is not found, a list containing the endpoints of an interval containing its value is returned.  Whether that interval is open, closed, or a mixed interval depends on the options passed; if the option {\tt Verbose} is set to {\tt true}, the precise interval will be printed.
